Classifier

### 2019-04-26 Code/dallin_annotations/07_build_classifier.Rmd
Explained the PCA and XGBoost aspects of our model.

---
title: "Classifier Construction"
output: github_document
---

@@

# Libraries
```{r}
library(tidyverse)
library(fastrtext)
library(magrittr) # for assignment operator %<>%  Dallin: I will try to remove the use of %<>% to comply with the tidyverese stylguide.
library(plotly)
require(randomForest)
library(Matrix)
source("05_define_clustering_functions.R")
source("06_define_prediction_functions.R")
# source("../complaint-prediction/R/logit_classifier_func.R") This file does not exist

library(xgboost)
library(caret)
library(car)

set.seed(123)
```

@@

@@

### Scree Plot of Principal Components
```{r}
# make a scree plot of the components
screeplot(train.pr, type = "line", npcs = 20)

# the first component must be the amount of information about a word learned by the fasttext model
# let's go with 15 components with this model.

@@

```{r}

save(train.pr, file = here:here("Output", "prcomp4.dat"))

save(train, file = here:here("Output", "train_data.dat"))

@@

load(here:here("Output", "train_data.dat"))

trainIndex <- createDataPartition(train$complaint_label, p = .8, 
                                  list = FALSE, 
                                  times = 1)

test_data <- train[-trainIndex,]
test_labels <- test_data %>% select(complaint_label) %>% transmute(label = complaint_label)
train_data <- train[trainIndex,]
train_labels <- train_data %>% select(complaint_label) %>% transmute(label = complaint_label)
train_data_vec <- train_data %>%  select(tweetFeatures, avg_vec_50, avg_vec_50_PCA) %>% unnest %>% mutate_all(as.numeric)
test_data_vec <- test_data %>%  select(tweetFeatures, avg_vec_50, avg_vec_50_PCA) %>% unnest %>% mutate_all(as.numeric)

dtrain <- xgb.DMatrix(sparse.model.matrix(~.-1, train_data_vec), label = as.numeric(train_labels$label))
dtest <- xgb.DMatrix(sparse.model.matrix(~.-1, test_data_vec), label = as.numeric(test_labels$label))
dtest1 <- xgb.DMatrix(sparse.model.matrix(~.-1, test_data_vec))
watchlist <- list(train = dtrain, eval = dtest)

param <- list(max_depth = 2, eta = 1, silent = 1, nthread = 2, 
              objective = logregobj, eval_metric = evalerror)

bst <- xgboost(data = sparse.model.matrix(~.-1, train_data_vec), label = as.numeric(train_labels$label), max.depth = 4,
               eta = 0.2, nthread = 2, nrounds = 100, objective = logregobj, eval_metric = evalerror)

test_data$prediction <- predict(bst, dtest1)

predict(bst, dtest1) %>% map_chr(function(x) ifelse(x > 0, "complaint", "non-complaint"))

test_data %<>%
  mutate(predict_label = case_when(prediction > 0 ~ "Complaint",
                                   TRUE ~ "Non-complaint"),
         complaint_label = case_when(complaint_label == 1 ~ "Complaint",
                                   TRUE ~ "Non-complaint")) %>% 
  mutate(accuracy = predict_label == complaint_label)

confusionMatrix(factor(test_data$complaint_label), factor(test_data$predict_label))

save(bst, file = here:here("Output", "xgb.modelPC4.dat"))
xgb.save(bst, here:here("Output", "xgb.modelPC4-xgb_version.dat"))
save(test_data, file = here::here("Output", "test_data.dat"))

@@

>>>>

---
title: "Classifier Construction"
output: github_document
allow_html_dependencies: true
---

@@

# Libraries
```{r}
library(tidyverse)
library(fastrtext)
library(magrittr) # for assignment operator %<>%  Dallin: I will try to remove the use of %<>% to comply with the tidyverese stylguide.
library(plotly)
require(randomForest)
library(Matrix)
library(xgboost)
library(caret)
library(car)
```

```{r}
source("05_define_clustering_functions.R")
source("06_define_prediction_functions.R")
# source("../complaint-prediction/R/logit_classifier_func.R") This file does not exist
```

```{r}
set.seed(123)
```

@@

# Dallin: This visualization of the PCA model was commented out because it caused problems with knitting.

@@

### Scree Plot of Principal Components
PCA models compress multi-dimensional issues into smaller "principal components." They identify trends and relationships in data, and these trends are summarized in "principal components." These principal components are responsible for describing a certain portion of the variance seen in the data.

We are using this scree plot to analyze how many principle components can summarize the variance in our data well. The scree plot shows that with only two principal components, a large portion of the variance in the data can be described, and little of the remaining variance remains.

When using our PCA model to make predictions, we want a model that is as simple as possible while describing the data well. In terms of PCA, this generally means that you want to use only a few principal components to describe the majority of variance in the data. The scree plot shown below helps do this by informing us how many principal components we should use to effectivly describe our training data.

The scree plot shows variances (in other words, variance undescribed by the model) on the y-axis, and principal components on the x-axis. For example, with only one principal component, a large amount of variance remains in the data. With two principal components, however, we see that the amount of undescribed variance drops considerably. The rule of thumb with the scree plot is to identify the "elbow" (where the slope of the line starts to flatten out), and choose a number of principal components based on number of principal components at that point. Using this plot, for example, we will use the first four principal components of the model. Using four principal components will help us achieve a model that is both simple, and accurate. Using more than four principal components will result in significantly diminished returns -- it will slow down our model without providing a significant improvement to our model's predictive ability. 
```{r}
# make a scree plot of the components
screeplot(train.pr, type = "line", npcs = 20)

# the first component must be the amount of information about a word learned by the fasttext model
# let's go with 15 components with this model.
# Dallin: Do we want to go with 15 components? Not sure why...

@@

### Output of VectorizeTweet() Function (PCA Class)
Each principal component can be thought of as a single, compressed dimension. The values shown below represent the magnitude of the tweet along that single dimension. Plotting these values should result in a plot similar to the 3D scatter plot shown above. 
```{r}
train$avg_vec_50_PCA[[1]]
```

# Save the Model Files
Here, we save our PCA model (prcomp4.dat) and our training dataset (train_data.dat -- a dataset containing tweet text, along with its respecive tweet vector and PCA values).

We then remove these models from the environment and reload them (this has helped ensure that our resulting output is truely independent.)
```{r}
save(train.pr, file = here::here("Output", "prcomp4.dat"))
save(train, file = here::here("Output", "train_data.dat"))

@@

load(here::here("Output", "train_data.dat"))
```

# Train the XGboost Model

### Create a Training Index
We need to separate our dataset (train_data.dat) into training data and test data for our xgboost model. The code below creates a list that randomly selects 80% of our dataset for training, leaving 20% of it for testing.
```{r}
trainIndex <- createDataPartition(train$complaint_label, p = .8, 
                                  list = FALSE, 
                                  times = 1)
```

### Group the Testing and Training Datasets
For each dataset (testing and training) we are creating a set of data containing the complaint labels associated with the tweet, as well as set containing the tweet's features, the tweet vector and the PCA output of the model.
```{r}
test_data <- train[-trainIndex,]
test_labels <- test_data %>% select(complaint_label) %>% transmute(label = complaint_label)
test_data_vec <- test_data %>%  select(tweetFeatures, avg_vec_50, avg_vec_50_PCA) %>% unnest %>% mutate_all(as.numeric)

train_data <- train[trainIndex,]
train_labels <- train_data %>% select(complaint_label) %>% transmute(label = complaint_label)
train_data_vec <- train_data %>%  select(tweetFeatures, avg_vec_50, avg_vec_50_PCA) %>% unnest %>% mutate_all(as.numeric)

@@

### Create Dense Matrices for XGBoost
XGBoost models need matrices as inputs, not dataframes or tibbles. Here, we take our testing and training data, and turn it into a matrix instead. To make it easy to keep track of our matrices, we have stored both the training and test matrices in a single list (called watchlist).

UPDATE: It looks like we don't even use these matrices later on..?
Nevermind. dtest1 is needed.
```{r}
# I don't think the lines I have commented out are actually necessary for the model 
# dtrain <- xgb.DMatrix(sparse.model.matrix(~.-1, train_data_vec), label = as.numeric(train_labels$label))
# dtest  <- xgb.DMatrix(sparse.model.matrix(~.-1, test_data_vec), label = as.numeric(test_labels$label))
dtest1 <- xgb.DMatrix(sparse.model.matrix(~.-1, test_data_vec))
# watchlist <- list(train = dtrain, eval = dtest)
```

### Set the Parameters of the Model and Begin Training

```{r}
# NOTE: All code below is commented out because this code had trouble knitting

# bst <- xgboost(data = sparse.model.matrix(~.-1, train_data_vec), label = as.numeric(train_labels$label), max.depth = 4,
#                eta = 0.2, nthread = 2, nrounds = 100, objective = logregobj, eval_metric = evalerror)
```

# Analyze the Model

### Get Predictions from the Model
Here, we let the model make predictions on our test dataset. We then compare the predictions to our actual labels to help us understand accuracy. 
```{r}
# test_data$prediction <- predict(bst, dtest1)
# 
# predict(bst, dtest1) %>% map_chr(function(x) ifelse(x > 0, "complaint", "non-complaint"))
#   
# 
# test_data <- test_data %>% 
#   mutate(predict_label = case_when(prediction > 0 ~ "Complaint",
#                                    TRUE ~ "Non-complaint"),
#          complaint_label = case_when(complaint_label == 1 ~ "Complaint",
#                                    TRUE ~ "Non-complaint")) %>% 
#   mutate(accuracy = predict_label == complaint_label)
```

### Look at the Confusion Matrix
This Confusion matrix helps us see how many false postives and false negatives we had. For example, there were 85 instances where a tweet was predicted as a complaint, but it was not (false positive), and 38 instances where a complaint was not predicted to be a complaint, when it actually was (false negative). Overall, the model had an 87.59% accuracy, which seems pretty good to me.  
```{r}
# confusionMatrix(factor(test_data$complaint_label), factor(test_data$predict_label))

# The document had trouble knitting after knitting, so I've copied and pasted the output of the code from the line above.

# Confusion Matrix and Statistics
# 
#                Reference
# Prediction      Complaint Non-complaint
#   Complaint           205            85
#   Non-complaint        38           663
#                                           
#                Accuracy : 0.8759          
#                  95% CI : (0.8537, 0.8958)
#     No Information Rate : 0.7548          
#     P-Value [Acc > NIR] : < 2.2e-16       
#                                           
#                   Kappa : 0.6852          
#                                           
#  Mcnemar's Test P-Value : 3.359e-05       
#                                           
#             Sensitivity : 0.8436          
#             Specificity : 0.8864          
#          Pos Pred Value : 0.7069          
#          Neg Pred Value : 0.9458          
#              Prevalence : 0.2452          
#          Detection Rate : 0.2069          
#    Detection Prevalence : 0.2926          
#       Balanced Accuracy : 0.8650          
#                                           
#        'Positive' Class : Complaint   
```

### Save the Model
```{r}
# save(bst, file = here::here("Output", "xgb.modelPC4.dat"))
# xgb.save(bst, here::here("Output", "xgb.modelPC4-xgb_version.dat"))
# save(test_data, file = here::here("Output", "test_data.dat"))
```

### 2019-04-26 Paper/paper.Rmd
Explained the PCA and XGBoost aspects of our model.

To use our model, we fed a set of tweets into a PCA model and a FastText classification model. The output of these two models is then fed into our XGBoost model, which makes a binary prediction for the tweet as a complaint or non-complaint.

@@

## Useful Resources:
Quick Explanation of PCA: https://www.youtube.com/watch?v=HMOI_lkzW08
Deeper Explanation of PCA: https://www.youtube.com/watch?v=_UVHneBUBW0

@@

## Useful Resources:
When I was first learning about XGBoost, I found these youtube videos to be helpful in explaining the very basic concepts behind boosted trees. Hopefully, they can help you too if you're struggling to follow along.

>>>>

To use our model, we fed a set of tweets into a PCA model and a FastText classification model. The output of these two models is then fed into our XGBoost model, which makes a binary prediction for the tweet as a complaint or non-complaint. I believe that the idea behind this "frankenstein model" is to use the perspectives of various styles of models to gain an accurate consensus on the data -- somewhat similar to the concept behind random forests.

@@

## Details of the Use of PCA in Our Model
We wanted to incorporate a PC analysis in our model, hoping that the perspective of a unique model type would aid our model's accuracy. As input, the PCA model needs numerical data, so we needed to convert our character data (tweet text) into numerical data. To do this, we used a fastText model again. Unlike the fastText model above, however, we decided to use an unsupervised fastText model for producing word vectors.

Ultimately, we used PCA because we wanted to take our text data and process it in a way it could become quantitative data, believing that this would be a better format as an input for our XGBoost model. It is important to note that the PCA model was not capable of predicting complaints vs non-complaints, and never "saw" training labels when we trained it (it was unsupervised). Rather, it processed the textual data to provide quantitative data that grouped complaints and non-complaints fairly well, despite being unaware of the labels.

### Using a fastText Unsupervised Model for Word Vectors
This model, trained on the tweet text from our training dataset of nearly 5,000 tweets, learns the associations of words by analyzing how frequently pairs of words appear near each other. After running this analysis, the model can produce a word vector for any word -- even words it hadn't seen before.

What are these word vectors? They represent a mathematical definition of a word. Via a method that is beyond the level of depth we'd like to explain in this paper, the output of this model (a vector containing 50 numbers for each word queried) is a representation of the associations one word has with all other words used in a corpus. Our model was specifically trained to produce word vectors of length 50, but this was simply a parameter we found to be most effective for our purposes (we explored this in the code fasttext_clust_EDA.R). Similar to the PCA model, it is important to note that this model that never "saw" our training labels (i.e. it was unsupervised), and was never capable of predicting complaints and non-complaints. Rather, it processed the textual data to provide a quantitative input for our principal component analysis.

These word vectors also function well with additions and subtractions. For example, with word vectors, king - man + woman = queen. This fact is very useful to us, because it means that if we average the vectors for each word in a tweet, we should end up with a summary of a tweet in a single vector. We created a function that averaged word vectors, named "VectorizeTweet()." The function takes a single tweet, breaks its text down into a dataframe of each word used in the tweet, fetches the word vector for each word in the dataframe, and then averages each column of the vectors. This vectorized tweet summary became the numerical input for our PC analysis.

### Analysis of Principal Components
Once we had a list-column of average tweet vectors (avg_vec_50), we analyzed its principal components. We started with a 3D visualization of the first three principal components, coloring the tweets by their complaint/non-complaint labels, and found that the complaints and non-complaints grouped well. We also looked at a scree plot and found that four principal components would be effective enough to describe the variance in our data well.

Once we saw that we should use four principal components, we then trained the PCA model using our dataset of about 5,000 tweets, and used the model to find the values for the four principal components for eqch tweet. This provided another numerical input we used in our next model, the XGboost model.

## Useful Resources for PCA:
Quick Explanation of PCA: https://www.youtube.com/watch?v=HMOI_lkzW08
Slightly Deeper Explanation of PCA: https://www.youtube.com/watch?v=_UVHneBUBW0

@@

## Our Use of the XGBoost Model
At this point, we had three sets of quantitative data on our tweets as inputs for our XGboost model: averaged tweet vectors, PCA outputs for four principal components, and a set of tweet features. I haven't mentioned the tweet features yet, but it was a simple step we took to gather crude data on the tweet. We created a function that calculated various details on a tweet, such as the number of spaces, the number of characters, and the number of exclamation points used in a tweet. After we ran this function on each tweet, the data was stored in a list-column called tweetFeatures.

We partitioned our dataset of 5,000 tweets (now processed to include tweet features, a tweet vector, and four principal component values) so that 80% was designated for training, and 20% was designated for testing. We then transformed our tibble of data into a dense matrix (the format required for XGBoost). We trained the XGBoost model, allowed the model to make predictions on our test dataset, and then compared the resulting predictions against our testing dataset's labels using a confusion matrix. We found our XGBoost model to be 87.59% accurate, with 85 false positives, and 38 false negatives.

## Useful Resources for XGBoost:
When I was first learning about XGBoost, I found these youtube videos to be helpful in explaining the very basic concepts behind boosted trees. Hopefully, they can help you too if you're struggling to follow along.
